### 2018-04-29 doc/paper.tex
Adding explanation of how K-compensated de Casteljau works.
The description is very raw and needs cleaning, but this is
a start.

Also re-formatted the "Example of Compensated de Casteljau Failing"
section to use `\mach` instead of `u` and to put a hat over the
computed values.

In this section we review the compensated de Casteljau algorithm
from \cite{Jiang2010}. In order to track the local errors at
each update step, we use four EFTs:
\begin{align}
\left[\widehat{r}, \rho\right] &= \mathtt{TwoSum}(1, -s) \\
\left[P_1, \pi_1\right] &= \mathtt{TwoProd}\left(\widehat{r}, \widehat{b}_j^{(k + 1)}\right) \\
\left[P_2, \pi_2\right] &= \mathtt{TwoProd}\left(s, \widehat{b}_{j + 1}^{(k + 1)}\right) \\
\left[\widehat{b}_j^{(k)}, \sigma_3\right] &= \mathtt{TwoSum}(P_1, P_2)
\end{align}
With these, we can exactly describe the local error between the exact
update and computed update:
\begin{gather}
\ell_j^{(k)} = \pi_1 + \pi_2 + \sigma_3 + \rho\widehat{b}_j^{(k + 1)} \\
(1 - s) \cdot \widehat{b}_j^{(k + 1)} +
  s \cdot \widehat{b}_{j + 1}^{(k + 1)} =
\widehat{b}_j^{(k)} + \ell_j^{(k)}.
\end{gather}

@@

In \texttt{CompDeCasteljau}, after a single stage of error filtering we
``give up'' and use \(\widehat{\partial b}^{(k)}\) instead of
\(\partial b^{(k)}\) (and don't keep around any information about the
round-off error). In order to obtain results that are as accurate as if
computed in \(K\) times the working precision, we must continue filtering
errors down \((K - 1)\) times, so that at the final level we compute
\(\widehat{\partial^{K - 1} b}^{(k)}\) and don't worry about round-off
error.

@@

Consider
\begin{equation}
p(s) = (4s - 3)^3 (8s + 7) = (-189) B_{0, 4}(s) + (-54) B_{1, 4}(s) +
57 B_{2, 4}(s) + (-32) B_{3, 4}(s) + 15 B_{4, 4}(s)
\end{equation}
at the point \(s = \frac{3}{4} + 800 u\). In exact arithmetic, we
have \(p(s) = 13(3200u)^3 + \mathcal{O}\left(u^4\right)\).
However, using the compensated de Casteljau algorithm results in
\(b_0^{(0)} + \partial b_0^{(0)} = 0\):

\begin{center}
  \begin{tabular}{>{$}c<{$} >{$}c<{$} >{$}c<{$} >{$}c<{$} >{$}c<{$} >{$}c<{$}}
    \toprule
    k & j & b_j^{(k)} & \partial b_j^{(k)} & \partial^2 b_j^{(k)} & \partial^3 b_j^{(k)} \\
    \midrule
    3 & 0 & -87\frac{3}{4} + 108032u & -32u & 0 & 0 \\
    3 & 1 & 29\frac{1}{4} + 88768u & 32u & 0 & 0 \\
    3 & 2 & -9\frac{3}{4} - 71200u & 0 & 0 & 0 \\
    3 & 3 & 3\frac{1}{4} + 37600u & 0 & 0 & 0 \\
    \midrule
    2 & 0 & 187200u & -15360000u^2 & 0 & 0 \\
    2 & 1 & -62408u & 8u - 128000000u^2 & 0 & 0 \\
    2 & 2 & 20800u & 87040000u^2 & 0 & 0 \\
    \midrule
    1 & 0 & -6u - 199688192u^2 & 6u - 99831808u^2 & -90112000000u^3 & 0 \\
    1 & 1 & -2u + 66568192u^2  & 2u + 33271808u^2 & 172032000000u^3 & 0 \\
    \midrule
    0 & 0 & -3u + 7296u^2 & 3u - 7296u^2 & 13 (3200u)^3 + 3048(512u)^4 & 962(128u)^4 \\
    \bottomrule
  \end{tabular}
\end{center}

>>>>

In this section we review the compensated de Casteljau algorithm
from \cite{Jiang2010}. In order to track the local errors at
each update step, we use four EFTs:
\begin{align}
\left[\widehat{r}, \rho\right] &= \mathtt{TwoSum}(1, -s) \\
\left[P_1, \pi_1\right] &= \mathtt{TwoProd}\left(\widehat{r}, \widehat{b}_j^{(k + 1)}\right) \\
\left[P_2, \pi_2\right] &= \mathtt{TwoProd}\left(s, \widehat{b}_{j + 1}^{(k + 1)}\right) \\
\left[\widehat{b}_j^{(k)}, \sigma_3\right] &= \mathtt{TwoSum}(P_1, P_2)
\end{align}
With these, we can exactly describe the local error between the exact
update and computed update:
\begin{gather}
\ell_j^{(k)} = \pi_1 + \pi_2 + \sigma_3 + \rho\widehat{b}_j^{(k + 1)} \label{ell-j} \\
(1 - s) \cdot \widehat{b}_j^{(k + 1)} +
  s \cdot \widehat{b}_{j + 1}^{(k + 1)} =
\widehat{b}_j^{(k)} + \ell_j^{(k)}.
\end{gather}

@@

In \texttt{CompDeCasteljau}, after a single stage of error filtering we
``give up'' and use \(\widehat{\partial b}^{(k)}\) instead of
\(\partial b^{(k)}\) (without keeping around any information about the
round-off error). In order to obtain results that are as accurate as if
computed in \(K\) times the working precision, we must continue filtering
errors down \((K - 1)\) times, so that at the final level we compute
\(\widehat{\partial^{K - 1} b}^{(k)}\) and don't worry about round-off
error.

When computing \(\widehat{\partial^F b}\) (i.e. the error after
\(F\) stages of filtering)
there will be several sources of round-off. In particular, there will be
\begin{itemize}
\item errors when computing \(\widehat{\ell}_{F, j}^{(k)}\) from the
  terms in \(\ell_{F, j}^{(k)}\)
\item an error
for the ``missing'' \(\rho \cdot \widehat{\partial^F b}_j^{(k + 1)}\) in
\((1 - s) \cdot \widehat{\partial^F b}_j^{(k + 1)}\)
\item an error from the product
  \(\widehat{r} \otimes \widehat{\partial^F b}_j^{(k + 1)}\)
\item an error from the product
  \(s \otimes \widehat{\partial^F b}_{j + 1}^{(k + 1)}\)
\item two errors from the two \(\oplus\) when combining the three
  terms in
  \(\left[\widehat{r} \otimes \widehat{\partial^F b}_j^{(k + 1)}\right] \oplus
  \left[s \otimes \widehat{\partial^F b}_{j + 1}^{(k + 1)}\right] \oplus
  \widehat{\ell}_{F, j}^{(k)}\)
\end{itemize}
For example, in~\eqref{ell-j}:
\begin{equation}
\ell_{1, j}^{(k)} = \underbrace{\pi_1}_{P_1 = \widehat{r} \otimes
    \widehat{b}_j^{(k + 1)}} +
\underbrace{\pi_2}_{P_2 = s \otimes \widehat{b}_{j + 1}^{(k + 1)}} +
\underbrace{\sigma_3}_{P_1 \oplus P_2} +
\underbrace{\rho \cdot \widehat{b}_j^{(k + 1)}}_{(1 - s)
  \widehat{b}_j^{(k + 1)}}
\end{equation}
After each stage, we'll always have
\[\ell_{F, j}^{(k)} = e_1 + e_2 + \cdots + e_M + \rho \cdot
\widehat{\partial^{F - 1} b}_j^{(k + 1)}\]
where the terms \(e_1, \ldots, e_M\) come from using \texttt{TwoSum} and
\texttt{TwoProd} when computing \(\widehat{\partial^{F - 1} b}_j^{(k)}\)
and the \(\rho\) term comes from the roundoff
in \(1 \ominus s\) when multiplying \((1 - s)\) by
\(\widehat{\partial^{F - 1} b}_j^{(k + 1)}\). With this in mind, we
can define an EFT (\texttt{LocalErrorEFT}) that computes
\(\widehat{\ell}\) and tracks all round-off errors generated in
the process:

\begin{algorithm}[H]
  \caption{\textit{EFT for computing the local error.}}

  \begin{algorithmic}
    \Function{\(\left[\eta, \widehat{\ell}\right] =
        \mathtt{LocalErrorEFT}\)}{$e, \rho, \delta b$}
      \State \(M = \texttt{length}(e)\)
      \\
      \State \(\left[\widehat{\ell}, \eta_1\right] =
          \mathtt{TwoSum}(e_1, e_2)\)
      \For{\(j = 3, \ldots, M\)}
        \State \(\left[\widehat{\ell}, \eta_{j - 1}\right] =
            \mathtt{TwoSum}\left(\widehat{\ell}, e_j\right)\)
      \EndFor
      \\
      \State \(\left[P, \eta_M\right] =
          \mathtt{TwoProd}\left(\rho, \delta b\right)\)
      \State \(\left[\widehat{\ell}, \eta_{M + 1}\right] =
          \mathtt{TwoSum}\left(\widehat{\ell}, P\right)\)
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\noindent With this EFT helper in place, we can perform \((K - 1)\)
error filtrations:

\begin{algorithm}[H]
  \caption{\(K\)-\textit{compensated de Casteljau algorithm.}}

  \begin{algorithmic}
    \Function{\(\mathtt{res} = \mathtt{CompDeCasteljau}\)}{$b, s$}
      \State \(n = \texttt{length}(b) - 1\)
      \State \(\left[\widehat{r}, \rho\right] = \mathtt{TwoSum}(1, -s)\)
      \\
      \For{\(j = 0, \ldots, n\)}
        \State \(\widehat{b}_j^{(n)} = b_j\)
        \For{\(F = 1, \ldots, K - 1\)}
          \State \(\widehat{\partial^F b}_j^{(n)} = 0\)
        \EndFor
      \EndFor
      \\
      \For{\(k = n - 1, \ldots, 0\)}
        \For{\(j = 0, \ldots, k\)}
          \State \(\left[P_1, \pi_1\right] = \mathtt{TwoProd}\left(
              \widehat{r}, \widehat{b}_j^{(k + 1)}\right)\)
          \State \(\left[P_2, \pi_2\right] = \mathtt{TwoProd}\left(
              s, \widehat{b}_{j + 1}^{(k + 1)}\right)\)
          \State \(\left[\widehat{b}_j^{(k)}, \sigma_3\right] =
              \mathtt{TwoSum}(P_1, P_2)\)
          \\
          \State \(e = \left[\pi_1, \pi_2, \sigma_3\right]\)
          \State \(\delta b = \widehat{b}_j^{(k + 1)}\)
          \\
          \For{\(F = 1, \ldots, K - 2\)}
            \State \(\left[\eta, \widehat{\ell}\right] =
                \mathtt{LocalErrorEFT}(e, \rho, \delta b)\)
            \State \(M = \texttt{length}(\eta)\)
            \\
            \State \(\left[P_1, \eta_{M + 1}\right] = \mathtt{TwoProd}\left(
                \widehat{r}, \widehat{\partial^F b}_j^{(k + 1)}\right)\)
            \State \(\left[P_2, \eta_{M + 2}\right] = \mathtt{TwoProd}\left(
                s, \widehat{\partial^F b}_{j + 1}^{(k + 1)}\right)\)
            \State \(\left[S_3, \eta_{M + 3}\right] =
                \mathtt{TwoSum}(P_1, P_2)\)
            \State \(\left[\widehat{\partial^F b}_j^{(k)}, \eta_{M + 4}\right]
                = \mathtt{TwoSum}\left(S_3, \widehat{\ell}\right)\)
            \\
            \State \(e = \eta\)
            \State \(\delta b = \widehat{\partial^F b}_j^{(k + 1)}\)
          \EndFor
          \\
          \State \(\widehat{\ell} =
                \mathtt{LocalError}(e, \rho, \delta b)\)
          \State \(\widehat{\partial^{K - 1} b}_j^{(k)} =
              \left(\widehat{r} \otimes
              \widehat{\partial^{K - 1} b}_j^{(k + 1)}\right) \oplus
              \left(s \otimes \widehat{\partial^{K - 1} b}_{j + 1}^{(k + 1)}
              \right) \oplus
              \widehat{\ell}\)
        \EndFor
      \EndFor
      \\
      \State \(\mathtt{res} = \widehat{b}_0^{(0)}\)
      \For{\(F = 1, \ldots, K - 1\)}
        \State \(\mathtt{res} = \mathtt{res} \oplus
            \widehat{\partial^F b}_0^{(0)}\)
      \EndFor
    \EndFunction
  \end{algorithmic}
\end{algorithm}

@@

Consider
\begin{equation}
p(s) = (4s - 3)^3 (8s + 7) = (-189) B_{0, 4}(s) + (-54) B_{1, 4}(s) +
57 B_{2, 4}(s) + (-32) B_{3, 4}(s) + 15 B_{4, 4}(s)
\end{equation}
at the point \(s = \frac{3}{4} + 800 \mach\). In exact arithmetic, we
have \(p(s) = 13(3200\mach)^3 + \mathcal{O}\left(\mach^4\right)\).
However, using the compensated de Casteljau algorithm results in
\(\widehat{b}_0^{(0)} \oplus \widehat{\partial b}_0^{(0)} = 0\):

\begin{center}
  \begin{tabular}{>{$}c<{$} >{$}c<{$} >{$}c<{$} >{$}c<{$} >{$}c<{$} >{$}c<{$}}
    \toprule
    k & j & \widehat{b}_j^{(k)} & \widehat{\partial b}_j^{(k)} & \widehat{\partial^2 b}_j^{(k)} & \widehat{\partial^3 b}_j^{(k)} \\
    \midrule
    3 & 0 & -87\frac{3}{4} + 108032\mach & -32\mach & 0 & 0 \\
    3 & 1 & 29\frac{1}{4} + 88768\mach & 32\mach & 0 & 0 \\
    3 & 2 & -9\frac{3}{4} - 71200\mach & 0 & 0 & 0 \\
    3 & 3 & 3\frac{1}{4} + 37600\mach & 0 & 0 & 0 \\
    \midrule
    2 & 0 & 187200\mach & -15360000\mach^2 & 0 & 0 \\
    2 & 1 & -62408\mach & 8\mach - 128000000\mach^2 & 0 & 0 \\
    2 & 2 & 20800\mach & 87040000\mach^2 & 0 & 0 \\
    \midrule
    1 & 0 & -6\mach - 199688192\mach^2 & 6\mach - 99831808\mach^2 & -90112000000\mach^3 & 0 \\
    1 & 1 & -2\mach + 66568192\mach^2  & 2\mach + 33271808\mach^2 & 172032000000\mach^3 & 0 \\
    \midrule
    0 & 0 & -3\mach + 7296\mach^2 & 3\mach - 7296\mach^2 & 13 (3200\mach)^3 + 3048(512\mach)^4 & 962(128\mach)^4 \\
    \bottomrule
  \end{tabular}
\end{center}
